---
title: "EDA"
author: "Jack Wright"
date: "4/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r}
library(tidyverse)
library(readxl)
train_data<-read_xlsx('./data/StudentData.xlsx')
```


## Data Structure

```{r}
glimpse(train_data)
```


There are 2571 observations of 32 predictors and the response (PH). All data are continuous besides the "Brand Code" which is unordered Categorical. 








## Exploring the Response


```{r}
train_data<-train_data%>%na.omit()
train_data%>%na.omit()%>%summarize(mean = mean(PH), "standard deviation" = sd(PH))
```


```{r}
ggplot(train_data, aes(x = PH))+geom_histogram()+ggtitle('Distribution of Response')
```




## Exploring the predictors






```{r}
library(dplyr)
library(psych)
library(car)
#library(asbio)
description<-psych::describe(train_data)
```


## Normality

```{r}
description%>%
  filter((abs(skew)>=.5))%>%nrow()

skew_list<-as.list(description%>%
  filter((abs(skew)>=.5))%>%rownames())
```


21 of the 32 numeric predictors are skewed.

skewed predictors can be found in the list `skew_list`

```{r}
ggplot(description, aes(x=range))+geom_histogram()+ggtitle('Ranges of Predictors')
```




Most of the predictors range is 1, but a minority are  much larger. Depending on the model this data will need to be `centered` and `scaled`



## Highly Correlated Predictor Pairs

```{r}
library(caret)
correlated_predictors<-findCorrelation(cor(train_data%>%select(-'Brand Code')), names = TRUE)

```

There are 6 highly correlated predictors and can be found in the `correlated_predictors` list



## PCA


```{r}
library(tidymodels)
pca_rec<-recipe(~., train_data%>%select(-'Brand Code')%>%select(-correlated_predictors))%>%
  update_role(PH, new_role = 'id')%>%
  step_normalize(all_predictors())%>%
  step_pca(all_predictors())
pca_prep<-prep(pca_rec)
```



```{r}

tidied_pca<-tidy(pca_prep, 2)

tidied_pca%>%
  filter(component %in% paste0('PC', 1:5))%>%
  mutate(component = fct_inorder(component))%>%
  ggplot(aes(value,terms,fill = terms))+
  geom_col(show.legend = FALSE)+
  facet_wrap(~component,nrow=1)+
  labs(y = NULL)

```


```{r}
library(tidytext)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  group_by(component) %>%
  top_n(8, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )
```

PCA 1 looks like it has a lot to do with pressure, as the positive elements are `Mnf Flow` `Hyd Pressure2` `Fill Pressure` and `Pressure Setpoint` , the negative elements are `Pressure Vaccum`and `Filler Level`. In this context positive means that as that variable increases we move in a positive direction along this principle component, while negative means that when that variable increases we move in the negative direction along that principle component. 

PCA 2 looks like it has a lot to do with the Carb, 



Now lets look at how these features are distributed in the plane of the first two Principle Components

```{r}
temp<-juice(pca_prep)%>%mutate(case_mean = case_when(PH > mean(PH)~TRUE, TRUE ~FALSE))
temp%>%
  ggplot(aes(PC1,PC2, label = PH, color = case_mean))+
  geom_point(alpha = .2)+
  geom_text(check_overlap = TRUE, family = 'IBMlexSans')+
  ggtitle('PH higher or lower than the mean along PC1 vs PC2')
```





There does seem to be some grouping in the PC1 vs PC2 plane, we could try some sort of `radial SVM` to model this. 







